### 2024-11-06 Report_Template_updated.Rmd
Initial commit

---
title: "Gillespie algorithm for simulating stochastic equations of molecular reactions"
author: "First  Last Name"
date: "`r Sys.Date()`" #  
output:
  pdf_document:
      number_sections: true
spacing: double
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{Title of Project}
- \fancyfoot[CO,CE]{}
- \fancyfoot[LE,RO]{\thepage}
- \usepackage{setspace}\doublespacing
---

```{=tex}
\newpage 
\tableofcontents 
\newpage
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Introduction
This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and 
MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

The rest of this project is as follows. In Section 2, we review some basic concepts. 
In Section 3, we introduce the topic. In Section 4, we have an application and simualtion. 
Section 5 provides conclusion and some closing remarks.

Section 2: original algorithm and basic theory
Section 3: modified algorithm
Section 4: example
Section 5: conclusion (how the modification is better, when to use)

# Review of the Topic

When you click the **Knit** button a document will be generated that includes both content as well 
as the output of any embedded R code chunks within the document. You can embed an R code chunk like 
this:

```{r cars,comment=NA}
summary(cars)
```

You can also embed publication quality plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Definitions and Examples
Such as importance of longitudinal data analysis (separate age and cohort effect like in the effect of age on reading ability), different correlation structures, missing data, etc.

## Details of the topic
This is a mathematical formua without formula number.
$$h(x) = \frac{f(x)}{S(x)} = -\frac{d \log[S(x])}{\rm dx}$$

This is how write a fomula and label it. 
\begin{align}
\label{eq2}
S(x) = \exp[-H(x)] = \exp[-\int_0^x h(u) {\rm d}u]
\end{align}

# Application
This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and 
MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

\begin{table}[!htp]
\small
\caption{This is a caption.}
\begin{center}
\begin{tabular}{|c|cc|}
\hline
& {Placebo}   &  {Treatment}       \\ \cline{1-3}  
Method& {Est} (SE)    &{Est} (SE) \\    %  95\% CI \hspace{0.5cm}  95\% CI 
\hline
\scriptsize{Method 1 } & ??? (???)     & ??? (???)      \\   %   (???, ???) \hspace{0.25cm} (???, ???)
\scriptsize{Method 2}  &  ??? (???)     &  ??? (???)     \\  %  (???,  ???)  \hspace{0.25cm}  (???, ???)
\hline
\end{tabular}
\label{tab: 2.3}
\end{center}
\end{table}

# Conclusion

# References
"Lu, Ting, et al. Cellular growth and division in the Gillespie algorithm. Systems biology 1.1 (2004):
121-128."