### 2024-11-06 Report_Template.Rmd
initial template

---
title: "Title of the Project"
author: "First  Last Name"
date: "`r Sys.Date()`" # or DEcember 1, 2023
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Introduction
This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and 
MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

The rest of this project is as follows. In Section 2, we review some basic concepts. 
In Section 3, we introduce the topic. In Section 4, we have an application and simualtion. 
Section 5 provides conclusion and some closing remarks.

# Review of the Topic

When you click the **Knit** button a document will be generated that includes both content as well 
as the output of any embedded R code chunks within the document. You can embed an R code chunk like 
this:

```{r cars,comment=NA}
summary(cars)
```

You can also embed publication quality plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Definitions and Examples
Such as importance of longitudinal data analysis (separate age and cohort effect like in the effect of age on reading ability), different correlation structures, missing data, etc.

## Details of the topic
This is a mathematical formua without formula number.
$$h(x) = \frac{f(x)}{S(x)} = -\frac{d \log[S(x])}{\rm dx}$$

This is how write a fomula and label it. 
\begin{align}
\label{eq2}
S(x) = \exp[-H(x)] = \exp[-\int_0^x h(u) {\rm d}u]
\end{align}

# Application
This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and 
MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

\begin{table}[!htp]
\small
\caption{This is a caption.}
\begin{center}
\begin{tabular}{|c|cc|}
\hline
& {Placebo}   &  {Treatment}       \\ \cline{1-3}  
Method& {Est} (SE)    &{Est} (SE) \\    %  95\% CI \hspace{0.5cm}  95\% CI 
\hline
\scriptsize{Method 1 } & ??? (???)     & ??? (???)      \\   %   (???, ???) \hspace{0.25cm} (???, ???)
\scriptsize{Method 2}  &  ??? (???)     &  ??? (???)     \\  %  (???,  ???)  \hspace{0.25cm}  (???, ???)
\hline
\end{tabular}
\label{tab: 2.3}
\end{center}
\end{table}

# Conclusion

# References