### 2024-11-26 test.Rmd
added rough observations to test.rmd

```{r simulation_1, echo=FALSE, warning=FALSE, message=FALSE}
#SLOW COMPUTATION.. FOR FAST COMPUTE SEE LINES 11, 20, 65, & 79 
# Load required libraries
if (!require("reshape2")) install.packages("reshape2")
if (!require("ggplot2")) install.packages("ggplot2")
library(reshape2)
library(ggplot2)
# Define reaction rates
c1 <- (37 / 1e6) * 10             # Formation of CPD
c2 <- (2 / 1e6) * 10              # Formation of 64-PP
c3 <- (1 - (37 + 2) / 1e6) * 10   # No dimerization occurs// REMOVE THIS LINE FOR FAST COMPUTE

# Initial state
state <- c(T1 = 10000, T2 = 10000, CPD = 0, `64PP` = 0)
# Define propensity functions
propensities <- function(state) {
  c(
    c1 * state["T1"] * state["T2"], # Reaction 1
    c2 * state["T1"] * state["T2"],  # Reaction 2
    c3 * state["T1"] * state["T2"]  #Reaction 3   // REMOVE THIS LINE FOR FAST COMPUTE
  )
}
# Gillespie simulation function
simulate_gillespie <- function(state, max_time) {
  # Initialize time and output data frame
  time <- 0
  output <- data.frame(
    Time = numeric(0),
    T1 = numeric(0),
    T2 = numeric(0),
    CPD = numeric(0),
    `64PP` = numeric(0)
  )

  # Record the initial state
  initial_row <- c(Time = time, state)
  output <- rbind(output, as.data.frame(t(initial_row), stringsAsFactors = FALSE))

  while (time < max_time) {
    # Calculate propensities
    props <- propensities(state)
    total_prop <- sum(props)

    # Stop if no reactions can occur
    if (total_prop == 0) break

    # Time to next reaction
    delta_t <- rexp(1, rate = total_prop)
    time <- time + delta_t

    # Select which reaction occurs
    reaction <- sample(length(props), 1, prob = props)

    # Update state based on reaction
    if (reaction == 1) {
      state["T1"] <- state["T1"] - 1
      state["T2"] <- state["T2"] - 1
      state["CPD"] <- state["CPD"] + 1
    }
    if (reaction == 2) {
      state["T1"] <- state["T1"] - 1
      state["T2"] <- state["T2"] - 1
      state["64PP"] <- state["64PP"] + 1
    }
    if (reaction == 3) {   #REMOVE THIS SECTION FOR QUICK COMPUTE
      next
    }

    # Record the new state
    new_row <- c(Time = time, state)
    output <- rbind(output, as.data.frame(t(new_row), stringsAsFactors = FALSE))
  }

  # Convert columns to numeric for consistency
  output[] <- lapply(output, as.numeric)
  return(output)
}
# Run the simulation
max_time <- 10        #SET TO HIGHER TIME IF DOING FAST COMPUTE
simulation <- simulate_gillespie(state, max_time)
```

```{r graph_1, echo=FALSE, warning=FALSE, message=FALSE}
# Reshape data for plotting
simulation_long <- melt(simulation, id.vars = "Time")
# Plot results
ggplot(simulation_long, aes(x = Time, y = value, color = variable)) +
  geom_line() +
  labs(
    title = "Dimerization Reaction Set Simulation",
    x = "Time",
    y = "Molecule Count",
    color = "Species"
  ) +
  theme_minimal()
```
```{r}
# Display the first few rows of the simulation data
knitr::kable(head(simulation), format = "html", caption = "Simulation Results (First Few Rows)")
```
###Observations
1.**Logarithmic Growth**
Both CPD and 64-PP formation follow logarithmic growth, as expected under specific UV radiation according to previously reported quantum yields (Martinez-Fernandez and Improta, 2022; Mouret et al., 2006). The number of reactions per unit time plateaus as the pyrimidine pairs decrease, consistent with the propensity functions.

2. **Stochastic Fluctuations**
Due to the stochastic nature of molecular reactions, this algorithm reflects fluctuations of photodimerization as it models the probabilistic process of each reaction.

3. **Time Steps and Reaction Events**

###Biological Significance
-**Photodimerization**
As pyrimidine photodimerization transforms two pyrimidine monomers into a dimer, this simulation has accurately represented the formation of two possible dimers, CPD and 64-PP, from pairs of thymine monomers. 
-**Real-World Application**
Dimerization is widely studied, as it can lead to DNA mutations causing skin cancer. 
This algorithm can be further adjusted to show molecular reactions with various molecular species and reaction channels.