### 2024-11-27 test.Rmd
new test for colouring links

---
title: "Test Document"
output:
  pdf_document:
    keep_tex: true
header-includes:
  - \usepackage{xcolor}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true, linkcolor=red, citecolor=green, urlcolor=blue}
---

This is a test document. See Figure \ref{fig:test}.

```{r fig-test, echo=FALSE, fig.cap="Test Figure \\label{fig:test}"}
plot(1:10)