### 2024-11-28 test.Rmd
added colours

---
title: "Test Document"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
    number_sections: true
    citation_package: biblatex
linkcolor: blue
spacing: double
header-includes:
  - \AtBeginDocument{\hypersetup{colorlinks=true,linkcolor=blue,filecolor=magenta,urlcolor=cyan}}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhead[L]{The Gillespie Algorithm}
  - \fancyfoot[CO,CE]{}
  - \fancyfoot[LE,RO]{\thepage}
  - \usepackage{setspace}
  - \doublespacing
  - \PassOptionsToPackage{hyphens}{url}  % Ensure URLs are broken correctly
  - \usepackage{xcolor} 
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true,linkcolor=blue,filecolor=magenta,urlcolor=cyan}
---

This is a test document. See Figure \ref{fig:test}.

```{r fig-test, echo=FALSE, fig.cap="Test Figure \\label{fig:test}"}
plot(1:10)